---
title: "proj"
author: "Yinglu Deng"
date: "10/18/2021"
output:
  html_document: default
  pdf_document: default
---

## Beijing PM2.5 Dataset Exploring

This hourly dataset contains the PM2.5 data of US Embassy in Beijing. Meanwhile, meteorological data from Beijing Capital International Airport are also included. (Source from https://archive-beta.ics.uci.edu/ml/datasets/beijing+pm2+5+data)

PM2.5 is a 2.5 micrometers or less than 2.5 particle pollutant exposure in the air. It increased hospital admissions for heart or lung causes, acute and chronic bronchitis.

### Attribute Information:
year: year of data in this row
month: month of data in this row
day: day of data in this row
hour: hour of data in this row
pm2.5: PM2.5 concentration (ug/m^3)
DEWP: Dew Point (℃)
TEMP: Temperature (℃)
PRES: Pressure (hPa)
cbwd: Combined wind direction
Iws: Cumulated wind speed (m/s)
Is: Cumulated hours of snow
Ir: Cumulated hours of rain


```{r cars}
# comment out after installation
#install.packages('raster')
#install.packages('ggmap')
#install.packages('viridis')
#install.packages('scales')

# load packages
library(ggmap)
library(raster)
library(scales)
library(viridis)
library(tidyverse) 
library(patchwork) 
library(dplyr)
library(MARSS)
library(forecast)

myData <- read.csv("PRSA_data_2010.1.1-2014.12.31.csv") 
head(myData)
#view(myData)

#completeness
missing_data_pm <- myData[!complete.cases(myData$pm2.5),]
round((dim(missing_data_pm)[1] / dim(myData)[1]) * 100, digit=0)

#basic descriptive statistics for each time series (at least mean, CV, ACF for each variable)
nrow(myData)   #rows in total
summary(myData)   # summarize data: min, mean, max, NAs

# create a column with month as a factor instead of integer
myData1 <- myData %>% 
  mutate(month_fac=factor(month)) # an ordered category
head(myData1)
#view(myData1)

# definition: The coefficient of variation (CV) is the ratio of the standard deviation to the mean. The higher the coefficient of variation, the greater the level of dispersion around the mean. It is generally expressed as a percentage.
CV <- function(x, ...){(sd(x, ...)/mean(x, ...))*100} # specify a function for CV

## calculate means and CVs with NAs removed using the na.rm=T argument
myData2 <- myData1 %>% 
  group_by(month_fac) %>% 
  summarize(pm2.5_mean=mean(pm2.5, na.rm=T), pm2.5_CV=CV(pm2.5, na.rm=T))
summary(myData2)
#view(myData2)

myData_day <- myData %>% 
  group_by(year, month, day) %>% 
  summarize(pm2.5_mean=mean(pm2.5, na.rm=T), pm2.5_CV=CV(pm2.5, na.rm=T))
summary(myData_day)
#view(myData_day)

date <-seq(lubridate::as_date("2010-1-1"), lubridate::as_date("2014-12-31"), by = "days")
#view(data)
myData_day <- myData_day %>% 
              add_column(date = date)
#view(myData_day)

# Plot an ACF of the ozone time series
#plot.ts(myData$pm2.5, main='PM2.5 Time Series', col = "purple", ylab = "PM2.5")

acf(myData$pm2.5, na.action = na.pass)
pacf(myData$pm2.5, na.action = na.pass)

p=myData_day%>%
  ggplot(aes(date,pm2.5_mean))+
  geom_line()+
  labs(title="Average PM2.5 in 2010-2014",x="Date", y = "Average PM2.5 (ug/m^3)")
print(p)
```
(1)The length of the time series is from the date 2010-01-01 to 2014-12-31. The frequency is one hour.
(2)There is 5% of NA's.
(3)There are total 43824 variables and 12 features in the dataset. The mean of pm2.5 is 98.61 and maximum is 994. After factor by month, the minimum pm2.5 mean is 80 and the maximum pm2.5 mean is 125.74. And the mean of pm2.5 coefficient of variation is 86.67. This autocorrelation represents the degree of positive correlation within 46 lags over successive time intervals. (Correlogram: [-1, 1])


# Q1: What is the day of average highest pm2.5 in each month?    
```{r code chunk 1}



month_day <-seq(lubridate::as_date("0000-1-1"), lubridate::as_date("0000-12-31"), by = "days")
myData_month_day <- myData %>% 
  group_by(month, day) %>% 
  summarize(pm2.5_mean=mean(pm2.5, na.rm=T), pm2.5_CV=CV(pm2.5, na.rm=T))
summary(myData_month_day)
myData_month_day <- myData_month_day %>% 
              add_column(month_day = month_day)
view(myData_month_day)

myData_day
Data_2014 <- filter(myData_day, year == 2014)
#view(myData_day)
#ggplot(myData_month_day, mapping = aes(x=month_day, y=pm2.5_mean)) + geom_point()
ggplot(Data_2014, mapping = aes(x=date, y=pm2.5_mean)) + geom_point()


p<-ggplot(Data_2014, aes(x=month, y=pm2.5_mean)) +
  geom_boxplot()
p
#p <- ggplot(myData_day, aes(x=pm2.5_mean, color=day)) +
#  geom_density()+
#  labs(title="PM2.5 Density Curve based on Season",x="PM2.5 concentration (ug/m^3)", y = "Density")
#p

# p <- ggplot(data=myData_day, aes(x=day, y=pm2.5_mean)) +
#  geom_bar(stat="identity")
# p 

# third week will be the worse of the the whole month?
#x-axis --year

```


# Q2.Which season has severer pm2.5 problem in a year?  (daliy)     auto.arima;forecast R package
```{r code chunk 2}
#view(myData)

myData$season[myData$month == 3 | myData$month == 4 | myData$month == 5] = "Spring"
myData$season[myData$month == 6 | myData$month == 7 | myData$month == 8] = "Summer"
myData$season[myData$month == 9 | myData$month == 10 | myData$month == 11] = "Autumn"
myData$season[myData$month == 12 | myData$month == 1 | myData$month == 2] = "Winter"

#view(myData)

myData_day2 <- myData %>% 
  group_by(season) %>% 
  summarize(pm2.5_mean=mean(pm2.5, na.rm=T), pm2.5_CV=CV(pm2.5, na.rm=T))
summary(myData_day2)



view(myData_day2)

#view(myData)

# winter mean is higher on the table
p <- ggplot(myData, aes(x=pm2.5, color=season)) +
  geom_density()+
  labs(title="Average PM2.5 Density Curve based on Season",x="Average PM2.5 concentration (ug/m^3)", y = "Density")
p

p<-ggplot(myData, aes(x=season, y=pm2.5, color=season)) +
  geom_boxplot()+
  labs(title="Box plot of Average PM2.5 based on Season",y="Average PM2.5 concentration (ug/m^3)", x = "Season")
p
# Add mean line
#p + geom_vline(data=myData_day2, aes(xintercept=pm2.5_mean),
 #           color=season, linetype="dashed")
```

# Q3.Does the problem of pm2.5 become more serious year by year? What is the trend?  get rid of hour (mean of day) predict 2015

```{r code chunk 3}
#group the data by year and month
myData_month <- myData %>% 
  group_by(year, month) %>% 
  summarize(pm2.5_mean=mean(pm2.5, na.rm=T), pm2.5_CV=CV(pm2.5, na.rm=T))
summary(myData_month)

#view(myData_month)

pm_mean <- myData_month[, 3, drop=FALSE] #take the pm2.5 average for each month
pm_mean_maxtric <- as.matrix(pm_mean) #transform to maxtric
auto.arima(pm_mean)  #time span: 2010-01-01 to 2014-12-31
pm_q3 <- ts(pm_mean, frequency = 12, start = c(2010, 1)) # let's convert it into a time series, otherwise auto.arima will not consider seasonal ARIMA models
auto.arima(pm_q3, seasonal=FALSE) # ARIMA(5,1,1)   non-seasonal
auto.arima(pm_q3, seasonal=TRUE) # ARIMA(2,1,2)(1,1,1)[12]   seasonal

#forecast one year into the future
pm_q3_model<-auto.arima(pm_q3, seasonal=TRUE)
pm_q3_forecast<-forecast(pm_q3_model, h=12)
plot(pm_q3_forecast,
  xlab="Year", ylab="Average PM2.5 (ug/m^3)") # plot forecast
```


# Q4.How does the PM2.5 work with the environment?  (MARSS with covariates)      
Null Hypothesis:
The environment has no effect on the PM2.5 (like Temperature, wind, snow, rain).

Alternative Hypothesis:
PM2.5 has effect on the environment.
(eg. higher pm2.5 concentration has higher temperature)

```{r code chunk 4}
myData_env <- myData %>% 
  group_by(year, month) %>% 
  summarize(pm2.5_mean=mean(pm2.5, na.rm=T), pm2.5_CV=CV(pm2.5, na.rm=T), 
            TEMP_mean=mean(TEMP, na.rm=T), Iws_mean=mean(Iws, na.rm=T), 
            Is_mean=mean(Is, na.rm=T), Ir_mean=mean(Ir, na.rm=T))
summary(myData_env)
#view(myData_env)

date <-seq(lubridate::as_date("2010-1-1"), lubridate::as_date("2014-12-31"), by = "month")
myData_env <- myData_env %>% 
              add_column(date = date)

#responses (or 'variates')
target <- cbind(myData_env$pm2.5_mean, myData_env$date)
target_maxtric <- as.matrix(target) #transform to maxtric
class(target_maxtric)
dim(target_maxtric)
target_maxtric

#view(myData_env)
#drivers (or 'covariates')
covar <- cbind(myData_env$TEMP_mean, myData_env$Iws_mean, myData_env$Is_mean, myData_env$Ir_mean)
covar_maxtric <- t(as.matrix(covar)) #transform to maxtric
covar_maxtric

covar1<-zscore(t(as.matrix(covar_maxtric[1,], drop=FALSE))) # 1st covariate
covar2<-zscore(t(as.matrix(covar_maxtric[2,], drop=FALSE))) # 2nd covariate
covar3<-zscore(t(as.matrix(covar_maxtric[3,], drop=FALSE))) # 3rd covariate
covar4<-zscore(t(as.matrix(covar_maxtric[4,], drop=FALSE))) # 4th covariate
zscore_covar_maxtric <- zscore(covar_maxtric)

dat <- zscore(t(target_maxtric)[1,])
dat

# Build model for covariate 1
mod0 = list()
# Define the parameters in the X equation (process model)
mod0$B = "identity"
mod0$U = "equal"  # you could compare to "equal"
mod0$Q = "diagonal and equal"
#The thing we are trying to get at is C, the effect of our driver on the species
mod0$C = "equal"
mod0$c = covar1 # This is the covariate 1 data
# Define the parameters in the Y equation (observation model)
mod0$Z = "identity"
  
# For R we set up a diag matrix with 0s where there are 0s and characters for things we want to estimate (one value per species)
mod0$R = "zero"
mod0$A = "zero"
  
# Let's fit the model
mod0.fit = MARSS(dat, model=mod0, control=list(maxit=1000))


# Do not bootstrap model unless you have a lot of time (it may take several hours)
kem.w.boot.CIs=MARSSparamCIs(mod0.fit,method="parametric",nboot=1000)  
kem.w.boot.CIs

# Build model for covariate 2
mod1 = mod0
mod1$c = covar2 # This is the covariate 1 data
# Let's fit the model
mod1.fit = MARSS(dat, model=mod1, control=list(maxit=1000))
kem.w.boot.CIs=MARSSparamCIs(mod1.fit,method="parametric",nboot=1000)  
kem.w.boot.CIs

# Build model for covariate 3
mod2 = mod0
mod2$c = covar3 # This is the covariate 1 data
# Let's fit the model
mod2.fit = MARSS(dat, model=mod2, control=list(maxit=1000))
kem.w.boot.CIs=MARSSparamCIs(mod2.fit,method="parametric",nboot=1000)  
kem.w.boot.CIs

# Build model for covariate 4
mod3 = mod0
mod3$c = covar4 # This is the covariate 1 data
# Let's fit the model
mod3.fit = MARSS(dat, model=mod3, control=list(maxit=1000))
kem.w.boot.CIs=MARSSparamCIs(mod3.fit,method="parametric",nboot=1000)  
kem.w.boot.CIs
```
List of covariates:
covariate 1: TEMP_mean -- mean Temperature (℃)
covariate 2: Iws_mean -- mean Cumulated wind speed (m/s)
covariate 3: Is_mean -- mean Cumulated hours of snow
covariate 4: Ir_mean -- mean Cumulated hours of rain

Target:
pm2.5: PM2.5 concentration (ug/m^3)

see the CI and if covers 0, means not significant -- null true
negative sign -- negative relationship between them
